---
title: "inLabZero: Basic Clustering"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{inLabZero: Basic Clustering}
  %\VignettePackage{inLabZero}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align="center",
  fig.width=6, 
  fig.height=6,
  out.width="50%",
  dpi=200,
  tidy=FALSE
)
options(width=80)
```

_version: 2020/01_


***

**_note_**: *This document looks into cluster analysis, as a part of the inLabZero package developed by
Inference Technologies. For more information on the inLabZero package, please refer to: [inLabZero introduction
vignette](InLabIntro.html)*

***


## Cluster Analysis 

 
Clustering (or cluster analysis) divides data into groups, called  **clusters**. Objects belonging to the same group are
in a sense more similar to each other than objects present in other groups (clusters). Cluster analysis can be achieved
by applying various algorithms, which use different approaches to cluster constitution, having also distinct efficacy in
cluster detection.


<br>

## Case Example: K-means Analysis 


**K-means** is an unsupervised machine learning algorithm that groups similar data points together while discovering
underlying patterns. In order to achieve this goal, it looks for a fixed number of clusters - **_k_** (_note_: in our
case example below, this corresponds to the parameter `centers`). This number is defined by the user and it refers to
the number of centroids the user estimates in the dataset (*_note_*: the real number of centroids is unknown before the
analysis itself). A centroid is the imaginary or real location representing the center of the cluster. Please note that
this parameter has to be chosen carefully. If we set the number of centroids too low, we can face a situation that we
omit a certain pattern (problem). A number that is too high might lead to pattern repetition. In such case, though, it
is possible to perform cluster manipulation and combine similar clusters into one (for more information on cluster
manipulation, please refer to [Cluster manipulation vignette](ClusterManipulation.html)).

K-means algorithm repeats 2 steps until convergence:

* Find the closest centroid to each point, and group points that share the same closest centroid 
* Update each centroid to be the mean of the points in its group


In this vignette, we will demonstrate two **applications of k-means analysis**:

1. [Performing k-means analysis on wafer bin maps with all bins](#allBins)
2. [Performing k-means analysis on wafer bin maps with single bin](#singleBin)

<br>

### Performing k-means analysis on wafer bin maps with all bins {#allBins}


The `model` function is a wrapper to perform regression, classification, clustering and dimensionality reduction. For
our example, we will define the type of function as `cluster` and `kmeans` is selected as the clustering model. The
model wrapper itself is run on wafer bin maps (`wbm`) dataset with data specification corresponding to models on maps.
Number of clusters, as previously brought up, is defined via the `centers` parameter.  

Clustering algorithm is applied to 100% of the data and the model outcome shows the number of wafers belonging to
particular clusters.

A detailed description of the models wrapper can be found in the corresponding [Modeling vignette](CART.html).


```{r}
library(inLabZero)
```

```{r echo=FALSE}
data(wbm)
```

```{r echo=FALSE}
mapPlotPar(default)
timePlotPar(default)
```

```{r}
# Perform clustering on all wafer bin maps applying the 'k-means' method with 6  
# user defined centroids
wbm <- model(wbm, data="map", type="cluster", model="kmeans", centers=6, seed=42)  
```


<br>

#### Viewing model summary 


By calling the summary function on the output of the model (stored in the previous step in the variable `wbm`), we can
view the model results summary. 

The summary contains all the clusters, as defined in the model function, summarizing number of wafers belonging to
particular clusters. Complete information on bins is also depicted. In the case of wafer bin maps, `bin0` occurrences
represent good dice. Any non zero bin represents failed die.


```{r}
summary(wbm)
```

<br>

#### Plotting a heatmap on all bins


To visualize clustering results let us plot the heatmaps (composite maps), where colors correspond to bad die frequency,
using the `mapPlot` plotting function. The input for the `heat` function represents the `bin` function, which in this
case creates 1 fail map, i.e. combines all the bins into `bin1`.

The figure below represents a composite map for a particular cluster.


```{r results="hide", fig.width=7, fig.height=10.5, out.width="100%"}
# Set the viewing settings for subsequent plotting via 'mapPlot' to 3 rows and 2 columns
par(mfrow=c(3,2))
mapPlot(heat(bin(wbm, "all", 1)))

```

<br>

#### Browse maps in cluster


Selection of wafers belonging to cluster `C2` is done via the `wbm[C2>0]` call (figure below depicts map corresponding
to row 1).

A detailed description of map browsing can be found in the corresponding [Map browsing vignette](MapBrowsing.html).


```{r eval=FALSE}
mapPlot(wbm[C2>0], numMain=T)
```
```{r echo=FALSE, results="hide"}
mapPlot(wbm[C2>0][1], numMain=T)
```



<br>

#### Model visualization: Time Series Plot


A plot of time series data can reveal a particular time span, during which the certain pattern occurred, which can
provide additional priceless information needed for the root cause analysis.


<br>

##### Displaying a time plot for a particular cluster


The following figure depicts a plot for the `ProbeYield` parameter, specifying cluster `C6` for `z` axis. This column
can be broken into intervals using `zLevel` (applicable for categorical variables). The data&time `x` axis is set to use
`probeEndData` variable. The main title can also be defined by the user and specified further in the `main` parameter.
The index of the plotted cluster is disabled for our use (`numMain=F`). In case we want to plot the points as an
equidistant sequence, we define `sequence=T`, then time intervals are not uniform. If we set this parameter to `FALSE`,
column `timeCol` will be used for the `x` axis. 


```{r}
# Set legend position to bottomright
timePlotPar(legend.pos="bottomright") 
# Plot time plot for `ProbeYield` selecting 'z' variable as 'C6' cluster with 
# 'probeEndDate' on the `x` axis plotting the points as an equidistant sequence
timePlot(wbm, ProbeYield, z="C6", zLevel="all", timeCol="probeEndDate", main="C6", 
			                                                       numMain=F, sequence=T)
```


<br>

##### Displaying multiple time plots for a series of clusters


Probe yield time series for all identified clusters can be seen in 6 time plots (grid of 3 rows and 2 columns).


```{r fig.width=7, fig.height=10.5, out.width="100%"}
# Set the viewing settings for subsequent plotting via 'mapPlot' to 3 rows and 2 columns
par(mfrow=c(3,2))
for (clust in 1:6) {
    cluster <- paste0("C", clust)
    timePlot(wbm, ProbeYield, z=cluster, zLevel="all", timeCol="probeEndDate",
    			                     main=cluster, numMain=F, sequence=T, showLegend=F)
}
```


### Performing k-means analysis on wafer bin maps with single bin {#singleBin}


In some cases it can be less noisy to perform analysis on wafer bin maps with single bin. Before the model is applied, a
single bin map is created. In the example below the `bin25` is selected.



```{r}
# Create single bin map
wbm1f <- bin(wbm, c(25))
# Perform clustering on all wafer bin maps with single bin applying the 'k-means' 
# method with 6 user defined centroids
set.seed(42)
wbm1f <- model(wbm1f, data="map", type="cluster", model="kmeans", centers=6)  
```


#### Viewing model summary 


By calling the summary function on the output of the model (stored in the previous step in the variable `wbm1f`), we can
view the model results summary for the single bin map.

The summary contains all the clusters, as defined in the model function, summarizing number of wafers belonging to
particular clusters. Information on the particular single bin is also depicted together with the good dice occurrences.
As mentioned above, in the case of wafer bin maps, `bin0` occurrences represent good dice. `bin1` represents failed dice (in this particular case, dice rejected in `bin25`).


```{r}
summary(wbm1f)
```

<br>

#### Plotting a heatmap on single bin wbm


To visualize clustering results let us plot the heatmaps, where colors correspond to bad die frequency,
using the `mapPlot` plotting function. In this case, we will plot the heatmap on single bin.

The figure below represents a composite map for a particular cluster.


```{r results="hide", fig.width=7, fig.height=10.5, out.width="100%"}
# Set the viewing settings for subsequent plotting via 'mapPlot' to 3 rows 
# and 2 columns
par(mfrow=c(3,2))
mapPlot(heat(wbm1f))

```


<br>

## Saving Maps


Saving data with clusters for future work is an important step, because clusters numbering, when clustering is repeated,
may change. Given the fact that the settings, usernames and file path/name are user and access specific, below example
is only informative. 


```{r eval=FALSE}
# illustrative example - user specific setting
saveData(wbm, "demo", user="demo", fileName="demo.all")
```

