---
title: "inLabZero: Cluster Manipulation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{inLabZero: Cluster Manipulation}
  %\VignettePackage{inLabZero}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align="center",
  fig.width=6, 
  fig.height=6,
  out.width="50%",
  dpi=200,
  tidy=FALSE
)
options(width=80)
```

_version: 2020/01_ 


***

**_note_**: *This document looks into cluster manipulation, as a part of the inLabZero package developed by Inference
Technologies. For more information on the inLabZero package, please refer to: [inLabZero introduction vignette](InLabIntro.html)*

***


## Cluster Manipulation 

Clustering (or cluster analysis) divides data into groups, called  **clusters**. Objects belonging to the same group are
in a sense more similar to each other than objects present in other groups (clusters). Cluster analysis can be achieved
by applying various algorithms, which use different approaches to cluster constitution, having also distinct efficacy in
cluster detection. Introduction into Cluster Analysis with a description of the K-means algorithm can be found in
[Basic Clustering vignette](BasicClustering.html). 

**Cluster manipulation** is an important part of the analysis. There are several possibilities how to manipulate with clusters: creation of new clusters, removing of existing clusters, moving/coping of specific wafers, lots or wafers belonging to a certain cluster to another cluster. 

During the analysis, when a certain wafer matches a pattern corresponding to a particular cluster, we can move this
wafer to the particular cluster. The source cluster is specified by the parameter `from` and the target cluster is
defined by the parameter `to`.


The logical parameter `add` of the `cluster` function enables moving or copying of selected wafers, lots or wafers from
a defined cluster into a cluster of choice. If `TRUE`, the selected group will be added (without alteration) to the
specified cluster. If `FALSE`, the selected group will be moved to the specified cluster and removed from the original
cluster. This parameter is set to `FALSE` by default.

<br>

### Example dataset 

During the cluster manipulation process, we will apply the cluster function to the wafer bin map dataset, `wbm`.
However, for the purposes of this vignette, we will show only a subset of the result output with columns relevant to the
topic e.g. the `lotWafer` parameter and the particular clusters that will be manipulated.

```{r}
library(inLabZero)
```
  
```{r echo=FALSE}
data(wbm)
```

```{r echo=FALSE}
# Subsetting only columns of interest for clarity
wbm <- wbm$tab[,c("lotWafer", "clear", "center", "edge")]
```

```{r}
wbm
``` 

```{r echo=FALSE}
data(wbm)
```

<br>

### Moving wafers to a particular cluster


We will at first demonstrate moving of all wafers belonging to a specific cluster to a different cluster. Since the
`add` parameter is by default `FALSE`, it is not necessary to specify it.

With the call below, we will move all wafers from cluster `clear` to cluster `myClust`, that will be created.

```{r results="hide"}
# Move all wafers from cluster 'clear' to cluster 'myClust'
moveWafers <- cluster(wbm, from=clear, to=myClust)
``` 

```{r echo=FALSE, results="hide"}
# Subsetting only columns of interest for clarity
moveWafers  <- moveWafers$tab[,c("lotWafer", "clear", "center", "edge", "myClust")]
```
```{r}
moveWafers
```

If we want to refine our selection and go to the wafer level, moving a particular wafer is also possible with the use of
the `lotWafer` parameter, which combines the information on the lot number and wafer number.

Let us move wafer number 8 from lot `LOT005` to cluster `center`.   

```{r results="hide"}

# Move wafer #8 from lot LOT005 to cluster center
moveWafer <- cluster(wbm, lotWafer=="LOT005-W08", to=center)
```

```{r echo=FALSE}
moveWafer  <- moveWafer$tab[,c("lotWafer", "clear", "center", "edge")]
```

```{r}
moveWafer
```

<br>

### Adding wafers and lots to a particular cluster 

In case we want to add wafers and lots to a particular cluster, we need to set the `add` parameter to `TRUE`. 

There are various possibilities of choosing the source to be added or moved. The example below shows how to add all
wafers from cluster `edge` to cluster `myClust`. In contrast to moving of wafers, there will not be any alteration to
the `edge` cluster. The newly created cluster, `myClust`, represents a copy of the cluster `edge`. 

```{r results="hide"}
# Add all wafers from cluster 'edge' to cluster 'myClust' 
addWafers <- cluster(wbm, from=edge, to=myClust, add=TRUE)
```

```{r echo=FALSE}
addWafers  <- addWafers$tab[,c("lotWafer", "clear", "center", "edge", "myClust")]
```

```{r}
addWafers
```

Now, instead of adding all wafers, we can also add particular lots to a specified cluster. As an example, we add lot
`LOT005` to cluster `clear`.  


```{r results="hide"}
# Add lot LOT005 to cluster clear
addLot <- cluster(wbm, lot=="LOT005", to=clear, add=TRUE)
```

```{r echo=FALSE}
addLot  <- addLot$tab[,c("lotWafer", "clear", "center", "edge")]
```

```{r}
addLot
```


<br>

### Creating a new cluster

The creation of a new cluster represents the analogy of adding wafers, lots or wafers belonging to a certain cluster to
a new user defined cluster. Since this corresponds to the action of 'adding', similarly as above, we need to specify the
`add` parameter as `TRUE`. 

In the first example, we create a new cluster with the name `newClust`, which will include the wafers from `LOT005`.  

```{r results="hide"}
# Create a new cluster, include lot LOT005
createCluster <- cluster(wbm, lot=="LOT005", to=newClust, add=TRUE)
``` 

```{r echo=FALSE}
createCluster  <- createCluster$tab[,c("lotWafer", "clear", "center", "edge", "newClust")]
```

```{r}
createCluster
```

The second example shows how we can define also various conditions to specify the level of data that we want to work
with. New cluster, `newClust2` will include data for only `part=PART1` with the probe End Date before `2016-03-05`. For
this particular example, we will include in the subset also `part` and `probeEndDate` information. As we can see,
although rows 1, 2 and 4 also correspond to `PART1`, the `probeEndDate` occured after `2016-03-05` and hence these do not
meet the condition and as a result, they are not added to `newClust2`. 

```{r results="hide"}
# Create a new cluster, include data only for `PART1` with probeEndDate before 5th March 2016
createCluster2 <- cluster(wbm, part=="PART1" & probeEndDate < "2016-03-05", 
                                                      to=newClust2, add=TRUE)
``` 

```{r echo=FALSE}
createCluster2  <- createCluster2$tab[,c("lotWafer", "part", "probeEndDate", 
                                      "clear", "center", "edge", "newClust2")]
```

```{r}
createCluster2
```

<br> 

### Remove a specific cluster

If we want to remove a certain cluster (e.g. one that was used temporarily during the analysis), we can specify the
cluster name for removal in the `remove` parameter of the function.

```{r echo=FALSE, results="hide"}
createCluster <- cluster(wbm, lot=="LOT005" & probeEndDate<"2016-03-05", 
                                                    to=newClust, add=TRUE)
```

```{r results="hide"}
# Remove cluster edge
removeCluster <- cluster(createCluster, remove="newClust")
```

```{r echo=FALSE}
removeCluster  <- removeCluster$tab[,c("lotWafer", "clear", "center", "edge")]
```

```{r}
removeCluster
```







