---
title: "inLabZero: Data Integration"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{inLabZero: Data Integration}
  %\VignettePackage{inLabZero}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align="center",
  fig.width=6, 
  fig.height=6,
  out.width="50%",
  dpi=200,
  tidy=FALSE
)
options(width=80)
```

_version: 2020/02_ 


***

**_note_**: *This document looks into data integration, as a part of the inLabZero package developed by Inference
Technologies. For more information on the inLabZero package, please refer to: [inLabZero introduction vignette](InLabIntro.html)*   

***

<br>

## Data integration

During the data analysis process, it is essential to be able to share data between different data layers (`pcy`, `wbm`, `leh`). In order to
obtain needed variables from diverse layers, proper data integration needs to be ensured.

The key function for variable transfer between dissimilar data layer objects represents the function `transfer`. The
first two parameters define the source (`from`) and the target object (`to`), respectively, followed by the column name
or names to be transferred. The parameter `filter` can be used for conditional transfer of variable data. This is
particularly useful in case we want to e.g. limit the yield information below a certain value. The aggregation function
can be specified through the `method` parameter, choosing from: `min`, `max`, `majority` (default), `mean` or `median`.
This method needs to be defined for transfer from higher to lower granularity. A crucial parameter applied during the
data integration represents the `key` parameter. This parameter specifies the merging key for the variable transfer
between different data layers. The default key will be chosen automatically as the highest intersect of granularities.
For `wbm`/`pcy` transfer, the `lotWafer` key is applied. For `wbm`/`leh`, `pcy`/`leh`, the `lot` key is used.


During the process of transfer between different data layers, it is of great importance to be aware of the various
granularities (or smallest units) of the datasets used. The table below summarizes this information in order to get a
more transparent view on this topic.   


##### Overview of minimum granularities for particular datasets


```{r table1, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- " 
| Dataset          | Granularity        | 1 row refers to | Additional information  | 
|-------------------|-------------------|-------------------|--------------------------------|
| pcy      | test site| 1 test site | 5 sites per wafer, 25 wafers per lot |  
| wbm      | wafer      | 1 wafer | 25 wafers per lot|  
| leh      | lot      |1 lot |     | 
"
cat(tabl) 
```

<br>

We can carry out the analysis at two granularities:

* wafer level
* site level

At **wafer level**, there is 1 value of yield per wafer (e.g. `ProbeYield`), which means that PCM parameters from all
test sites on the wafer are eventually combined with the same yield value.  At **site level**, the yield is calculated
in the vicinity of the test site and this `PCMYield` is combined only with PCM parameters from this particular test
site. With the site level approach, we can significantly reduce information noise and it is possible to pick up only
selected segments of wafers for analysis (using the filter option). If there is a need to utilize the whole wafer yield,
then wafer level is preferred.

In principle, we can also do the analysis on the lot level, but this procedure leads to a sizable data reduction.
Therefore it is suitable for issue investigation that mostly affects the whole lot. In this case, the rules for transfer
to lower granularity apply. 

In this vignette, we will demonstrate the following **applications of data integration**:

1. [Data integration - wafer level](#waferLevel) <br>
    1.1 [Transfer to higher granularity *(wbm to pcy, leh to wbm, leh to pcy)*](#wl_to_H) <br>
    1.2 [Transfer to lower granularity *(pcy to wbm, wbm to leh, pcy to leh)*](#wl_to_L)

2. [Data integration - site level](#siteLevel) <br>
    2.1 [Transfer to higher granularity *(wbm to pcy, leh to pcy)*](#sl_to_H)

<br>

***

**_note_**: During the data integration process, we will apply the `transfer` function to the wafer bin map dataset,
`wbm`, the process control dataset, `pcy`, and the lot equipment history dataset, `leh`. However, for the purposes of
this vignette, we will show only a subset of the result output with columns relevant to the topic e.g. the `lotWafer`
parameter and the particular variables that are of interest.  

***



<br> 

## 1. Data integration - wafer level {#waferLevel}

Let us concentrate on the approach, when the die yield per wafer represents the lowest yield resolution.

<br>

### 1.1 Transfer to higher granularity {#wl_to_H}


When considering transfer from lower to higher granularity, we have the following alternatives:




##### Transfer options to higher granularity 


``` {r table2, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- " 
|                           |test site          |  | wafer            | | lot                |
|---------------------------|-------------------|--|------------------|--|-------------------|
| wafer level to site level |pcy                |<-|wbm               |  |                   |
| lot level to wafer level  |                   | | wbm              |<-| leh               |
| lot level to site level   |pcy                | |                  |<-| leh               |
"
cat(tabl) 
```


<br>

####  Wafer level to site level 

In the case of wafer level to site level transfer, we are transferring a particular variable from the wafer bin map
dataset, `wbm`, to the process control data set, `pcy`.  

For demonstration purposes, let us create a new cluster, `newClust`, in the `wbm` dataset that will be subsequently used
for transferring. This cluster will include all the wafers from cluster `edge`. 

For more information on cluster manipulation, please refer to [Cluster manipulation vignette](ClusterManipulation.html).


```{r}
library(inLabZero)
```

```{r echo=FALSE}
data(wbm)
data(pcy)
data(leh)
```

```{r echo=FALSE}
mapPlotPar(default)
sitePlotPar(default)
```

```{r results="hide"}
# Create a new cluster, add all wafers from cluster edge to cluster newClust
wbm <- cluster(wbm, from=edge, to=newClust, add=TRUE)

```

```{r echo=FALSE}
# Subsetting only columns of interest for clarity
wbm <- wbm$tab[,c("lotWafer", "wafer", "part", "UPYield", "ProbeYield", 
  "clear", "center", "edge", "newClust")]
```

Looking at our wafer bin map dataset, we can see that the new cluster was added to `wbm`. 

```{r}
wbm
```

```{r echo=FALSE, results="hide"}
data(wbm)
wbm <- cluster(wbm, from=edge, to=newClust, add=TRUE)
```

Now we can apply the `transfer` function and store the result in our new variable, `pcyc`, which will contain the `pcy`
dataset along with the newly transferred column: `newClust`. The following call can be applied: 

```{r results="hide"}
# Transfer cluster 'newClust' from wbm to pcy
pcyc <- transfer(wbm, pcy, "newClust")
```

```{r echo=FALSE}
# Subsetting only columns of interest for clarity
pcyc <- pcyc$tab[,c("lotWaferSiteid", "lotWafer", "wafer", "part", "UPYield", 
  "ProbeYield", "siteid", "newClust", "PARAM01", "PARAM02", "PARAM03")]
```

Our process control dataset, stored in `pcyc`, contains new column `newClust`, that was previously only part of the
`wbm` dataset. In order to demonstrate the process of transfer to higher granularity, we will subset the `pcyc` and
focus on two wafers (Wafer 12 from `LOT001` and wafer 18 from `LOT165`), one of which belongs to cluster `newClust`.
Each wafer contains 5 rows for 5 different sideIDs. All siteIDs have 1 `ProbeYield` value. 


```{r}
pcyc[lotWafer %in% c("LOT001-W12", "LOT165-W18")]
```

```{r echo=FALSE}
# Subsetting only columns of interest for clarity
wbm <- wbm$tab[,c("lotWafer", "wafer", "part", "UPYield", "ProbeYield", "newClust" )]
```

Now we can see the point of the transfer to higher granularity: 1 value of `ProbeYield` that previously corresponded to
1 wafer in the `wbm` dataset is now matched after the transfer in `pcyc` to all siteIDs and hence to five rows.



```{r echo=FALSE, results="hide"}
data(wbm)
wbm <- cluster(wbm, from=edge, to=newClust, add=TRUE)
```

<br>

##### Transfer limited by filter 

As described above, conditional transfer of variable data can be achieved through the use of the `filter` parameter. The
call, `filter=(pcy[,ProbeYield]>96`, limits the transferred data (`newClust`) to `ProbeYield` values above 96%.  



```{r results="hide"}
pcycf <- transfer(wbm, pcy, "newClust", filter=(pcy[,ProbeYield]>96))   
```

```{r echo=FALSE}
# Subsetting only columns of interest for clarity
pcycf <- pcycf$tab[,c("lotWaferSiteid", "lotWafer", "wafer", "part", "UPYield", 
  "ProbeYield", "siteid", "newClust", "PARAM01", "PARAM02", "PARAM03")]
```

Process control dataset after data integration with `newClust` column complying with the set conditions is shown below.
Wafer number 12 from lot `LOT001` that previously belonged to `newClust` does not meet the `ProbeYield` condition and
thus after this transfer is not present in the `newClust` cluster.

```{r}
pcycf
```

<br>

##### Number of wafers after integration  

The information on the number of wafers present after the integration is obtainable in the following way:

```{r}
length(unique(pcycf[newClust>0, lotWafer]))
``` 

The notion `newClust>0` is applied for wafer presence check in the particular cluster (column), where presence
is equal to 1. The filter used in the above section applied strict rules and therefore, in our case,
there are no wafers present in the cluster `newClust` with `ProbeYield` above 96%. 

Looking at the `pcyc` variable after the first transfer, where `newClust` represents a copy of the `edge` cluster, we
see that 158 wafers are now present in the `newClust` column.

```{r}
length(unique(pcyc[newClust>0, lotWafer]))
``` 

<br>

#### Lot level to wafer level  

If we are interested in equipment history data corresponding to a particular process stage, we can transfer this
information from the lot equipment history dataset, `leh`. In this example, we choose to transfer stage `STAGE008` from
the `leh` dataset into the `wbm` dataset. 


```{r results="hide"}
wbm <- transfer(leh, wbm, "STAGE008") 
```

```{r echo=FALSE}
# Subsetting only columns of interest for clarity
wbm <- wbm$tab[,c("lotWafer", "lot", "wafer", "part", "UPYield", "ProbeYield",
 "newClust", "STAGE008")]
```

Wafer bin map dataset now contains `STAGE008` after data integration: 

```{r}
wbm
```

We can explain the transfer to higher granularity (`leh` -> `wbm`) on lot `LOT131`. The `leh` dataset (shown below)
lists equipment `EQ025` for `STAGE08`, which is presented on 1 row. 


```{r echo=FALSE}
leh <- leh$tab[,c("lot", "part", "STAGE008")]
# Subsetting only columns of interest for clarity
```


```{r}
leh[lot=="LOT131"]
```

After the transfer, this information is matched to every wafer from `LOT131` (contains 18 wafers altogether). 

```{r}
wbm[lot=="LOT131"]
```

<br>

Using the `mapPlot` function, we can plot the corresponding map for `LOT001`, `wafer 12`. The use of the `subMain`
parameter enables to specify the particular information to be plotted below the map. In our case, it is the equipment
name used at `STAGE008`. 


```{r echo=FALSE, results="hide"}
data(wbm)
data(leh)
wbm <- transfer(leh, wbm, "STAGE008") 
```

```{r}
mapPlot(wbm[lotWafer=="LOT001-W12"], subMain="STAGE008")
```

<br>
  
#### Lot level to site level 

Similar situation arises, when transferring a process stage of interest into the process control dataset, `pcy`.   

```{r echo=FALSE}
data(pcy)
data(leh)
```

```{r results="hide"}
pcy <- transfer(leh, pcy, "STAGE008") 
```

In order to explain the transfer to higher granularity (`leh` -> `pcy`), let us focus on lot `LOT005`. The `leh` dataset
lists equipment `EQ012` for `STAGE08` (presented on 1 row).

```{r echo=FALSE}
lehc <- leh$tab[,c("lot", "part", "STAGE008")]
# Subsetting only columns of interest for clarity
```

```{r}
lehc[lot=="LOT005"]
```

```{r echo=FALSE, results="hide"}
# Subsetting only columns of interest for clarity
pcy <- pcy$tab[,c("lotWaferSiteid", "lotWafer", "part", "UPYield", "ProbeYield", 
  "siteid", "PARAM01", "PARAM02", "PARAM03", "STAGE008")]
```

After the transfer, this information is matched to every siteID from `LOT005`.

```{r}
pcy[lotWafer=="LOT005-W08"]
```


```{r echo=FALSE}
data(pcy)
data(wbm)
```


<br>

### 1.2 Transfer to lower granularity {#wl_to_L}

Transferring from higher to lower granularity is summarized in the table below.

##### Transfer options to lower granularity 


``` {r table3, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- " 
|                           |test site          |  | wafer              |  |  lot               |
|---------------------------|-------------------|--| -------------------|--| -------------------|
| site level to wafer level |pcy                |->| wbm                |  |                    |
| wafer level to lot level  |                   |  | wbm                |->|  leh               |
| site level to lot level   |pcy                |->|                    |  |  leh               |
"
cat(tabl) 
```

During the transfer to lower granularity, we need to define the method that will be applied for aggregation. The
particular aggregation functions (specified via the `method` parameter) are: `min`, `max`, `majority` (default), `mean`
or `median`. The necessity of the aggregation is related to different granularities. This means that we have 5 times
more rows in the `pcy` dataset than in the `wbm` dataset. When transferring from `pcy` to `leh`, we are assigning 25x5
rows to 1 lot. If we choose `mean` as the aggregation function in this case, we will use the mean of 25x5 values for the
assignment.

##### Number of rows for `pcy` and `wbm`: 

```{r}
nrow(pcy)
nrow(wbm)
```


<br>

#### Site level to wafer level 

If we want to add a certain parameter that is present in the `pcy` dataset to the `wbm` dataset, we can use the function
calls as per below. In this case, we are transferring the parameter `PARAM10` using `mean` as the aggregation function. 

```{r results="hide"}
wbm <- transfer(pcy, wbm, "PARAM10", method="mean")
```
In the `pcy` dataset, `PARAM10` is available for 5 different sideIDs (5 rows for a particular wafer).

```{r echo=FALSE}
# Subsetting only columns of interest for clarity
data(pcy)
pcy <- pcy$tab[,c("lotWaferSiteid", "lotWafer", "wafer", "part", "UPYield", "ProbeYield", 
                  "siteid", "PARAM10")]
```


```{r}
pcy[lotWafer=="LOT001-W12"]
```

In the `wbm` dataset, 1 row corresponds to 1 particular wafer (in this example, we are focusing on wafer 12 from
`LOT001`). Using the aggregation function, `mean`, the value of `PARAM10` after the transfer to `wbm` will represent the
mean value of the 5 values listed above in the `pcy` dataset.


```{r echo=FALSE}
# Subsetting only columns of interest for clarity
wbm <- wbm$tab[,c("lotWafer", "wafer", "part", "UPYield", "ProbeYield", 
  "clear", "center", "edge", "PARAM10")]
```

```{r}
wbm
``` 

<br>

Using the `mapPlot` function, we can plot the corresponding map for `LOT001`, `wafer 12`. This time we will choose
`PARAM10` to be plotted below the map, via the the `subMain` parameter.


```{r echo=FALSE, results="hide"}
data(wbm)
data(leh)
data(pcy)
wbm <- transfer(pcy, wbm, "PARAM10", method="mean")
```

```{r results="hide"}
mapPlot(wbm[lotWafer=="LOT001-W12"], subMain="PARAM10")
```


```{r echo=FALSE}
data(wbm)
```

<br>

#### Wafer level to lot level 

Imagine we want to have the information on particular clusters in the lot equipment history dataset, `leh`. Using the
examples below, we can transfer the `clear` and `edge` columns to `leh` dataset using `mean` as the aggregation
function. 


```{r results="hide"}
lehc <- transfer(wbm, leh, "clear", method="mean")  
lehc <- transfer(wbm, lehc, "edge", method="mean") 
``` 




```{r echo=FALSE}
# Subsetting only columns of interest for clarity
wbm <- wbm$tab[,c("lotWafer", "lot", "wafer", "part", "UPYield", "ProbeYield", 
  "clear", "center", "edge")]
```

In this example, we will look at `LOT005`, which  contains 6 wafers and clusters `clear`, `center` and `edge`.

```{r}
wbm[lot=="LOT005"]
```

```{r echo=FALSE}
# Subsetting only columns of interest for clarity
lehc <- lehc$tab[,c("lot", "part", "clear", "edge", "STAGE001", "STAGE002", "STAGE003")]
```

As a result of this transfer process, we see that clusters `clear` and `edge` are now contained in the lot equipment
history dataset, `lehc` (below). We were aggregating each row representing one wafer from `LOT005` to 1 row representing
one lot, using the mean function.

```{r}
lehc[lot=="LOT005"]
``` 

After the integration, `lehc` contains information on the presence of a cluster in the particular lot with values in the
range from 0 to 1. Since we have chosen the `mean` function for aggregation, these values represent a mean value of
wafer presence with the cluster in the lot. As a result, we will only obtain 1 in case that all wafers in the given lot
have that cluster. Generally, this number represents a decimal number, which denotes the percentage of wafers that have
the cluster present in the lot. 

Therefore, the following cases are of interest for aggregation of clusters:

* `mean` - percentage of wafers with the cluster present in the lot
* `min` - equals 0, if at least one wafer in the lot does not have the cluster, otherwise equals 1
* `max` - equals 1, if at least one wafer in the lot has the cluster, otherwise equals 0

<br>

```{r echo=FALSE}
data(leh)
data(wbm)
data(pcy)
```

<br>

#### Site level to lot level 

Similarly, transferring a particular parameter from the `pcy` dataset to the `leh` dataset, we apply the following call.
`Median` is used as the aggregation function, in this case.


```{r results="hide"}
leh <- transfer(pcy, leh, "PARAM10", method="median")  
```


```{r echo=FALSE}
# Subsetting only columns of interest for clarity
data(pcy)
pcy <- pcy$tab[,c("lotWaferSiteid", "lotWafer", "wafer", "part", "UPYield", "ProbeYield", 
                  "siteid", "PARAM10")]
``` 

We will focus on wafer 12 from `LOT001` In the `pcy` dataset. `PARAM10` is listed for 5 different sideIDs (5 rows for a
particular wafer).

```{r}
pcy[lotWafer=="LOT001-W12"]
```


```{r echo=FALSE}
# Subsetting only columns of interest for clarity
leh <- leh$tab[,c("lot", "part", "STAGE001", "STAGE002", "STAGE003", "PARAM10")]
```
Looking at `leh` dataset after the transfer, we can see that it includes now `PARAM10` column, where each row contains
the median of `PARAM10` values from all test sites in particular lot (shown above).

```{r}
leh[lot=="LOT001"]
```



<br>

## 2. Data integration - site level {#siteLevel}


### PCMYield map 
    
Another approach consists in looking at the yield calculated in the test site vicinity. This yield is denoted as
`PCMYield`. Each wafer contains 5 test sites. The greenish square represents the area around the test site (small white
square in the middle), where the `PCMYield` is calculated. Each designated area around the test site is therefore
represented by 1 value of `PCMYield`. The value of the mean is shown in the bottom part of the squares. 

```{r echo=FALSE}
data(wbm)
data(pcy)
data(leh)
```

In order to plot the `PCMYield` parameter's distribution center per site, we can apply the `sitePlot` function. The
centering function applied on `PCMYield` data can be specified through the `fun` parameter. In our example, we use the
`mean` function, therefore we will plot the `PCMYield` mean of all wafers in the `pcy` dataset.   


```{r}
sitePlot(pcy, PCMYield, fun=mean)
```




<br>

### 2.1 Transfer to higher granularity {#sl_to_H}

In the previous section, where we discuss data integration on wafer level, we have demonstrated all the transfer
combinations between data levels. In this section, we will further look at some specific cases of transfer to higher
granularity related to test site and `PCMYield`.
 


##### Transfer options to higher granularity 

``` {r table4, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- " 
|                           |test site          |  | wafer            | | lot                |
|---------------------------|-------------------|--|------------------|--|-------------------|
| wafer level to site level |pcy                |<-|wbm               |  |                   |
| lot level to site level   |pcy                | |                  |<-| leh               |
"
cat(tabl) 
```


<br>

####  Wafer level to site level 

In the case of wafer level to site level transfer, we are transferring a particular variable from the wafer bin map
dataset, `wbm`, to the process control data set, `pcy`.  


```{r echo=FALSE, results="hide"}
data(wbm)
wbm <- cluster(wbm, from=edge, to=newClust, add=TRUE)
```

We will again apply the `transfer` function and store the result in our new variable, `pcyc`, which will contain the
`pcy` dataset along with the newly transferred column: `newClust`.  We can transfer variables conditionally using the
`filter` parameter. Filter `filter=(pcy[,PCMYield]<95` limits the transferred `newClust` data to `PCMYield` below 95%. 

The following call can be applied:

```{r results="hide"}
pcyc <- transfer(wbm, pcy, "newClust", filter=(pcy[,PCMYield]<95))   
```

```{r echo=FALSE}
# Subsetting only columns of interest for clarity
pcyc <- pcyc$tab[,c("lotWafer", "part", "siteid", "UPYield", "PCMYield", 
  "clear", "center", "edge", "newClust", "PARAM01", "PARAM02", "PARAM03")]
```

After the transfer, our process control dataset, stored in `pcyc`, contains new column `newClust` that was previously
only part of the `wbm` dataset. By applying the filter to the `PCMYield`, we can see that the `newClust` is not
transferred to siteid #3, which is the wafer center.  


```{r}
pcyc
```


<br>

##### Number of wafers per site after integration  

The information on the number of wafers per test site present after the integration is obtainable in the following way:

```{r echo=FALSE, results="hide"}
data(wbm)
data(pcy)
data(leh)
wbm <- cluster(wbm, from=edge, to=newClust, add=TRUE)
pcyc <- transfer(wbm, pcy, "newClust", filter=(pcy[,PCMYield]<95))   
```


```{r}
table(pcyc[,newClust], pcyc[,siteid])   
``` 


<br>

##### PCMYield map for selected cluster 

We can use again the `sitePlot` function to plot the wafer map of the mean `PCMYield` yield per test site while
requesting `pcyc[newClust>0]`.


```{r}
sitePlot(pcyc[newClust>0], PCMYield, fun=mean)
```

<br>
      
#### Lot level to site level  

Similar situation arises, when transferring a process stage of interest into the process control dataset, `pcy`. 


```{r echo=FALSE}
data(pcy)
```

```{r results="hide"}
pcy <- transfer(leh, pcy, "STAGE008") 
```

In order to explain the transfer to higher granularity (`leh` -> `pcy`), let us focus on lot `LOT005`. The `leh` dataset lists
equipment `EQ012` for `STAGE08` (presented on 1 row).

```{r echo=FALSE}
leh <- leh$tab[,c("lot", "part", "STAGE008")]
```

```{r}
leh[lot=="LOT005"]
```


```{r echo=FALSE, results="hide"}

# Subsetting only columns of interest for clarity
pcy <- pcy$tab[,c("lotWaferSiteid", "lotWafer", "part", "UPYield", "ProbeYield", 
                  "siteid", "PARAM01", "PARAM02", "PARAM03", "STAGE008")]
```
After the transfer, this information is matched to every siteID from `LOT005`.

```{r}
pcy[lotWafer=="LOT005-W08"]
```






