---
title: "inLabZero: Hist plot"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{inLabZero: Hist plot}
  %\VignettePackage{inLabZero}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}   
---

```{r, include = FALSE} 
knitr::opts_chunk$set( 
  collapse = TRUE,
  comment = "#>",
  fig.align="center",
  fig.width=6, 
  fig.height=6,
  out.width="50%",
  dpi=200,
  tidy=FALSE
)
options(width=80)   
```

_version: 2020/05_    


***

**_note_**: *This document looks into Hist plot, as a part of the inLabZero package developed by Inference Technologies. For more  information on the inLabZero package, please refer to: [inLabZero introduction vignette](InLabIntro.html)*             
 
***


## Data visualization    

Data visualization refers to presenting data in a graphical form. It is a unique tool for assessing relationships between attributes, spotting outliers or potential issues, evaluation of model performance and can even provide an idea for possible data transformations that need to be applied.  

In this vignette, the following inLabZero **topics related to Hist plot visualization** are documented:  

1. [Hist plot - `histPlot`](#section1) 
2. [Hist plot visual parameters - `histPlotPar`](#section2)  


<br>

### 1. Hist plot - `histPlot` {#section1} 

Histogram is a column-chart which depicts the underlying frequency of a set of discrete or continuous data that are measured on an interval scale. Thus, each column represents a range of the values and the height of a column corresponds to how many values there are in that range. It is a powerful visualization tool to easily inspect the underlying distribution of the data set and investigate other properties, such as skewness and kurtosis.

A histogram is created using the inLabZero function `histPlot`, where chart colors are based on `z`- axis variable. The function can be applied on the process control dataset, `pcy` (corresponding to parameter `obj`). The parameters for plotting are specified either as variable indexes (*integer*) or as variable names (*character*).  

<br>

#### Specification of the input parameter for plotting  

Let us look on two different ways to specify parameters with the same result output (1 graph will be shown for demonstration). The first corresponds to the case, where the parameters to be plotted are specified via a *variable index* (10th variable), the second uses a *variable name* for parameter specification (non-quoted strings can be used).

```{r}
library(inLabZero)  
```

```{r echo=FALSE}
data(pcy)
```

```{r echo=FALSE}
histPlotPar(default)
modelPlotPar(default)
```

```{r eval=FALSE}
# Plot 10th variable from the group
histPlot(pcy, 10)
```

```{r}
# Plot PARAM10 variable
histPlot(pcy, PARAM10)
```

<br> 
<br> 

It is also possible to plot a sequence of variables. In this case browsing mode is entered. For help, type `h` during browsing. 

For example (don't include the quotation marks):

<ul>
<li><strong>[enter]</strong>: next</li>
<li><strong>"5"</strong>: go next 5 steps</li>
<li><strong>"-"</strong>: go one step back</li>
<li><strong>"-3"</strong>: go 3 steps back</li>
<li><strong>"auto"</strong>: autobrowsing</li>
<li><strong>"x"</strong>: exit</li>
</ul>

```{r eval=FALSE}
# Plot 10th to 15th variable from the group
histPlot(pcy, c(10:15))
```


<br>

##### numMain

Parameter **`numMain`** enables to include/exclude index of the plotted variable in the main title. It is set to `TRUE` by default, the parameter index is excluded in this example.

```{r}
# Plot PARAM10 variable
histPlot(pcy, PARAM10, numMain=F)
```

<br> 

##### outlier

An important part of data analysis consists in outlier filtering. Parameter **`outlier`** is applied for filtering of variable before plotting. Outliers are data points outside the interval: `<(Q1 - kl * IQR), (Q3 + kr * IQR)>`, where `Q1` and `Q3` are quartiles, `IQR` is interquartile range and `k1` and `k2` are values specified by outlier parameter as its first and second element. If single value is provided, `k1` = `k2` setting is used. No outlier filtering is applied by default. 


```{r}
# Plot PARAM10 variable with outlier filtering 
histPlot(pcy, PARAM10, outlier=3, numMain=F)
```

<br> 
<br> 

The next couple of examples are divided into two sections based on whether the `z`- axis is numerical or categorical along with the usage of several other `histPlot` arguments selected based on the `z`- axis type.

<br>

#### `histPlot` argument settings for numerical `z`- axis

Parameter **`z`** specifies the variable for the `z` axis. For a numerical `z`- axis, this variable can be broken into intervals. Subsets of variable corresponding to `z` variable breaks are then plotted with different colors.

```{r}
# Select numerical z variable with input by name
histPlot(pcy, PARAM10, z="PCMYield", numMain=F)
```

<br> 

##### zBr

Parameter **`zBr`** breaks of a numeric `z` variable. Single integer value determines number of breaks. This parameter can also be specified as a vector and it can be numerical, character or date in a POSIXct class, then it represents interval breaking points. It is also possible to use special character values `min`, `mean`, `median` and `max`. 

```{r}
# Break numerical variables into intervals: keywords seq, min, max together with values
histPlot(pcy, PARAM10, z="PCMYield", zBr=c("min",seq(10,90,15),"max"), numMain=F)
```

<br> 

##### zBrMethod

The **`zBrMethod`** parameter defines the split methods and it works in conjunction with the `zBr` setting. The following split methods can be used:
**`median`** (breaks `z` variable with `median`), **`mean`** (breaks `z` variables with `mean`), **`thr`** (breaks `z` variable with threshold value specified by `zBr`), **`yieldHeat`** (corresponds to `zBr=c("seq(min,median,length.out=brValue)","max")`, where `brValue` is the value specified by `zBr`), **`linear`** (breaks `z` variable into equidistant intervals; number of splits is given by `zBr`). The following call uses the `linear` option as the `zBrMethod`: 

```{r}
# Plot PARAM10 histogram with 'linear' option as the zBrMethod
histPlot(pcy, PARAM10, z="PARAM14", zBr=10, zBrMethod="linear", numMain=F)
```

<br> 

##### zOutlier

In order to filter out outliers corresponding to the `z` variable, that would distort the real range of the variable, the `zOutlier` parameter can be applied. Parameter **`zOutlier`** is an adaptation of the `outlier` parameter for outliers filtering of the `z` variable before plotting.   

To *combine multiple plots into one overall graph* we can use function **`par()`** and include the option **`mfrow=c(nrows, ncols)`**. 

```{r fig.height=10.5}
par(mfrow=c(2,1)) 
histPlotPar(legend.pos="topright")
# Plot PARAM15 histogram with input by name
histPlot(pcy, PARAM15, z="PARAM77", zBr=5, zBrMethod="linear", numMain=F)
# Plot PARAM15 histogram with outlier filtering option
histPlot(pcy, PARAM15, z="PARAM77", zBr=5, zBrMethod="linear", zOutlier=3, numMain=F) 
```

```{r echo=FALSE}
par(mfrow=c(1,1))   
```

```{r echo=FALSE}
histPlotPar(default)
```

<br>

##### Date type `z` variable

The following example shows the use of the `zBr` parameter specified as POSIXct class to demonstrate date type `z` variable.

```{r}
# Define date type z variable
histPlot(pcy, 10, z="pcStartDate", 
                zBr=as.POSIXct(c("2015/10/01", "2016/06/01", "2016/10/01")))
```


<br>

#### `histPlot` argument settings for categorical `z`- axis


Categorical `z`- axis can be broken into intervals using the **`zLevel`** parameter (*note*: In order to *include all levels*, use special value `zLevel="all"`). Subsets of variable corresponding to `z` variable breaks are then plotted with different colors. For a categorical `z`- axis, the `zLevel` parameter is non-null. It specifies which categories to use for split. The examples below can be compared with regards to the order of input for the `zLevel` parameter and the overlay depiction in the graphs.   

```{r}
# Select categorical z variable with input by name and define z-axis breaking intervals
histPlot(pcy, PARAM10, z="center", zLevel=c(0,1), numMain=F)
```

```{r}
# Reverse drawing order
histPlot(pcy, PARAM10, z="center", zLevel=c(1,0), numMain=F)
```

<br>

#### Chart types


Parameter **`chart`** specifies chart type. We can choose from the following values:

* `hist` - plots histogram layers
* `density` - plots density layers (set as default)
* `beside` - plots histogram groups for each bin


<br> 

##### Histogram layers plot

With the `hist` option for the parameter `chart`, we will plot histogram layers. 

```{r}
# Select categorical z variable with input by name and define z-axis breaking intervals
histPlot(pcy, PARAM10, chart="hist", z="center", zLevel=c(0,1), numMain=F)
```

<br>

##### xBr 

The **`xBr`** parameter breaks of a numeric `x` variable. Single integer value determines number of breaks. This parameter can also be specified as a vector and it can be numerical or character. 
 

<br>

##### xBrMethod

The **`xBrMethod`** parameter defines the split methods and it works in conjunction with the `xBr` setting. The following split methods can be used:
**`median`** (breaks `x` variable with `median`), **`mean`** (breaks `x` variables with `mean`), **`yieldHeat`** (corresponds to `xBr=c("seq(min,median,length.out=brValue)","max")`, where `brValue` is the value specified by `xBr`), **`linear`** (breaks `x` variable into equidistant intervals; number of splits is given by `xBr`). The following call uses the `linear` option as the `xBrMethod`: 


```{r}
# Plot PARAM10 histogram with 'linear' option as the xBrMethod
histPlot(pcy, PARAM10, chart="hist", xBr=10, xBrMethod="linear", z="center", zLevel=c(0,1))
```


<br>

##### xLevel 

Categorical `x`- axis can be broken into intervals using the **`xLevel`** parameter (*note*: In order to *include all levels*, use special value `xLevel="all"`). For a categorical `x`- axis, the `xLevel` parameter is non-null. It specifies which categories to use for split.


```{r}
# Select categorical z variable with input by name and define x-axis breaking intervals
histPlot(pcy, wafer, xLevel=c(1:10), chart="hist", z="center", zLevel="all") 

```

<br>
 

##### Histogram groups plot for each bin

The x-axis interval of a histogram is divided into several equal parts, which are denoted as **bins**. Let us now plot histogram groups for each bin, setting the `chart` parameter to `beside`.  

```{r}
# Select categorical z variable with input by name and define z-axis breaking intervals
histPlot(pcy, PARAM10, chart="beside", z="center", zLevel=c(0,1), numMain=F)
```

<br> 

In order to break the `x` - axis into equidistant intervals, we use the `linear` option for the `xBrMethod` in conjunction with the `xBr` parameter, which determines the number of breaks. 

```{r}
# Plot PARAM10 histogram with 'linear' option as the xBrMethod
histPlot(pcy, PARAM10, chart="beside", xBr=8, xBrMethod="linear", z="center", zLevel=c(0,1))
```

<br> 

##### yLim

There is also an option to set `y` axis limits via the **`yLim`** parameter. It represents a vector with two elements corresponding to the minimum and the maximum value of the axis. The following call shows how to set the user defined `y` axis limits:


```{r}
# User defined y axis limits
histPlot(pcy, PARAM10, chart="beside", yLim=c(0,20), z="center", zLevel="all", numMain=F)
```

<br> 

Next, we will show two options for layout tuning.

<br>

##### main

The main title can be added via the **`main`** parameter.

```{r eval=FALSE}
# Add graph title
histPlot(pcy, PARAM10, z="center", zLevel=c(0,1), 
          main="Center pattern generated at low PARAM10", showLegend=T)
```
<br>

##### showLegend


To disable the displaying of the legend, set the **`showLegend`** parameter to `FALSE`.


```{r}
# Add title and hide legend
histPlot(pcy, PARAM10, z="center", zLevel=c(0,1), 
         main="Center pattern generated at low PARAM10", showLegend=F)
```


<br>

#### Visualization of the model output

The limits identified by the CART model could be verified in the histogram. For more details regarding the CART model see the [CART vignette](CART.html).  

In this example, we use the value of 2 to balance the train set and select cluster `clear` for negative case sampling.  

```{r results="hide"}
 # Train CART classification on balanced data
pcyb <- model(pcy, type="class", model="cart", formula="center ~ .", cp=0.08, balance=2, 
                                                              balanceCol="clear", seed=42)  
```

Model visualizations with the use of the `modelPlot` function:

```{r}
modelPlot(pcyb, varlen=0, main="CART, cluster, classification, prune, balanced", 
                                                                        extra=108)      
``` 

<br> 

In our analysis, we are working with wafers in the `center` cluster and we are randomly sampling twice the amount of the wafers from the cluster `clear` for balancing. In case of repeating the balancing process, random sampling will vary. 

To ensure that we use for the histogram same balanced data as for the model, we will store the selected wafers in the `usedRow` variable using the `modelTrainRow` function. The `usedRow` variable will now contain all the wafers with the `center` cluster and samples from the cluster `clear`.   

```{r}
usedRow <- modelTrainRow(pcyb)
```

Since we have stored all the wafers in the `center` cluster and samples from cluster `clear` in the `usedRow` variable, the `histPlot` can be called on `pcyb[usedRow]` subset so that only the data used in the model is displayed. The CART limits are added using `abline` function, a red horizontal line at the value of 1713.  


```{r}
histPlotPar(legend.pos="topleft")
 # Select categorical z variable with input by name for all levels
histPlot(pcyb[usedRow], PARAM10, z="center", zLevel="all", numMain=F)
# add CART limit - red vertical line
abline(v=1713, col="red", lty=3)  
```

<br>

### 2. Hist plot visual parameters - `histPlotPar` {#section2}  

The **`histPlotPar`** function enables to get, set and reload default visual parameters of the `histPlot`. For a detailed list of all parameters that are available, see `?histPlotPar`. Use parameter names to get values, use assignment to set parameter values and use keyword `default` to load default values. Let us look at some of the most used parameters and their usage. In order to see the effect on the `histPlot`, the `histPlotPar` needs to be run before the `histPlot` is plotted. 


<br>

##### axes.x.signif

To get the value of a visual parameter, in our case value of significant digits in x axis, use: 

```{r}
# Get value of significant digits in x axis
histPlotPar(axes.x.signif)
```

<br>

##### legend.pos

The *legend position* is accessible via the **`legend.pos`** parameter. In the call below, the legend position is set to top left: 

```{r}
# Set legend position to top left     
histPlotPar(legend.pos="topleft")
histPlot(pcy, PARAM10, z="center", zLevel=c(0,1), numMain=F)
```

```{r echo=FALSE}
histPlotPar(default)
```

<br> 

##### colors.zLevel

*Color scheme* for categorical variables can be tuned via the **`colors.zLevel`** parameter. Please refer to [color palette](https://colorbrewer2.org/#type=qualitative&scheme=Set3&n=12) for more details on color specification.  

```{r}
# Set color scheme to user defined values
histPlotPar(colors.zLevel=c("#8dd3c7","#fdb462"), legend.pos="topright")
histPlot(pcy, PARAM10, z="center", zLevel=c(0,1), numMain=F) 
```

```{r echo=FALSE}
histPlotPar(default)
```

<br> 

##### colors.alpha.density


Parameter **`colors.alpha.density`** represents the Alpha setting (transparency) for density layers' chart.  

```{r}
# Set the Alpha setting (transparency) for density layers' chart
histPlotPar(colors.alpha.density=0.8)
histPlot(pcy, PARAM10, z="center", zLevel=c(0,1), numMain=F)
```

<br> 

In case that we have changed the visual parameters and we want to reload the default settings, the following call can be applied: 

```{r}
# Reload default visual parameters
histPlotPar(default) 
```




