---
title: "inLabZero: Map Browsing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{inLabZero: Map Browsing}
  %\VignettePackage{inLabZero}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align="center",
  fig.width=6, 
  fig.height=6,
  out.width="50%",
  dpi=200,
  tidy=FALSE
)
options(width=80)
```

_version: 2020/01_


***

**_note_**: *This document presents various ways of map browsing, as a part of the inLabZero package developed by
Inference Technologies. For more information on the inLabZero package, please refer to: [inLabZero introduction
vignette](InLabIntro.html)*   

***


## Various ways of map browsing

During the issue identification process, we can make use of plots that give us better insight into the problem and
pattern localization helps to identify the origin of the underlying problem.

There are various ways to tweak the map plot viewing options, which will be demonstrated below.



<br>

### Map Browser

In order to view a particular map, we call the `mapPlot` function on the wafer bin map dataset, as demonstrated below. 

While displaying more than 1 map, the browsing mode is entered. For help, input `[h]` during browsing. 

The *browser options* comprise (don't include the quotation marks):

<ul>
<li><strong>[enter]</strong>: next</li>
<li><strong>"5"</strong>: go next 5 steps</li>
<li><strong>"-"</strong>: go one step back</li>
<li><strong>"-3"</strong>: go 3 steps back</li>
<li><strong>"auto"</strong>: autobrowsing</li>
<li><strong>"x"</strong>: exit</li>
</ul>


```{r}
library(inLabZero)
```

```{r eval=FALSE}
mapPlot(wbm)
```

```{r echo=FALSE}
mapPlotPar(default)
```

```{r echo=FALSE, results="hide"}
data(wbm)

mapPlot(wbm[1], numMain=T)
```



<br>

### Specifying (multiple) maps for viewing


Using the `wbm[1:100]` call, we can specify multiple maps for viewing. This command plots wafer maps  corresponding to
rows 1 to 100, in the browsing mode (figure below depicts map corresponding to row 1). 

Parameter `numMain` enables to include index of the plotted row in the main title.


```{r, eval=FALSE}
mapPlot(wbm[1:100], numMain=T)
```

```{r, echo=FALSE, results="hide"}
mapPlot(wbm[1], numMain=T)
```


<br>

### Map browsing in a lot

It can be useful to browse through maps belonging to a specific lot. This can be specified by using the
`wbm[lot=="LOT005"]` call (in this case, maps corresponding to lot `LOT005` are selected for browsing).

```{r eval=FALSE}
mapPlot(wbm[lot=="LOT005"]) 
```
```{r echo=FALSE, results="hide"}
mapPlot(wbm[lot=="LOT005"][1]) 
```


<br>

### Selection of map by lot and wafer number


It is possible to select wafer map by lotWafer parameter `lotWafer=="LOT005-W08"` (combination of the lot and wafer
number). 

The use of the `subMain` parameter enables to specify the particular information to be plotted below the map. In our
case, it is the probe end date.


```{r eval=FALSE}
mapPlot(wbm[lotWafer=="LOT005-W08"], subMain="probeEndDate") 
```
```{r echo=FALSE, results="hide"}
mapPlot(wbm[lotWafer=="LOT005-W08"][1], subMain="probeEndDate") 
```


<br>

### Map browsing in a cluster


Selection of wafers belonging to cluster `edge` is done via the `wbm[edge>0]` call (figure below depicts map
corresponding to row 1).


```{r eval=FALSE}
mapPlot(wbm[edge>0], numMain=T)
```
```{r echo=FALSE, results="hide"}
mapPlot(wbm[edge>0][1], numMain=T)
```


<br>

### Comparison of a single bin map in a cluster to an all bin map

Once we have created a single bin map, via `wbm1f <- bin(wbm, c(25))`, we can compare it to an all bin map, for a
particular cluster of interest. In this case, for the single bin map on the left, `bin1` represents failed dice (here:
dice rejected in `bin25`). 


```{r eval=FALSE}
# Create single bin map
wbm1f <- bin(wbm, c(25))
mapPlot(wbm1f[edge>0], wbm)

```
```{r echo=FALSE, results="hide", fig.width=14, fig.height=7, out.width="100%"}
wbm1f <- bin(wbm, c(25))
mapPlot(wbm1f[edge>0][1], wbm[1])
```

<br>

### Editing maps in a cluster

The function option `edit=TRUE` enables to edit maps in a particular cluster. During this mode, the user can add wafer
to the cluster, move wafer from particular cluster to another cluster or delete wafer from all clusters.


Available *browser options* in this mode:

<ul>
<li><strong>add cluster:</strong> add a specific wafer to a particular cluster </li>
<li><strong>move cluster:</strong> move a specific wafer to a particular cluster</li>
<li><strong>del:</strong> delete wafers from all clusters</li>
</ul>

```{r eval=FALSE}
mapPlot(wbm, cluster=edge, edit=TRUE)
```
```{r echo=FALSE, results="hide"}
mapPlot(wbm[1],  cluster=edge, edit=TRUE)
```




