---
title: "inLabZero: Time plot"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{inLabZero: Time plot}
  %\VignettePackage{inLabZero}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---

```{r, include = FALSE} 
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align="center",
  fig.width=6, 
  fig.height=6,
  out.width="50%",
  dpi=200,
  tidy=FALSE
)
options(width=80) 
```

_version: 2020/04_  


***

**_note_**: *This document looks into Time plot, as a part of the inLabZero package developed by Inference Technologies. For more  information on the inLabZero package, please refer to: [inLabZero introduction vignette](InLabIntro.html)*        
 
***


## Data visualization 

Data visualization refers to presenting data in a graphical form. It is a unique tool for assessing relationships between attributes, spotting outliers or potential issues, evaluation of model performance and can even provide an idea for possible data transformations that need to be applied.  

In this vignette, the following inLabZero **topics related to Time plot visualization** are documented:

1. [Time plot - `timePlot`](#section1) 
2. [Time plot visual parameters - `timePlotPar`](#section2) 


<br>

### 1. Time plot - `timePlot` {#section1} 


The **time plot** is used to render the variable as a time series, which is a set of data points or observations taken at specified times, usually at the same intervals (e.g. hourly, daily, weekly, etc.).

A time plot is created using the inLabZero function `timePlot`, where chart colors are based on `z`- axis variable. The function can be applied on the process control dataset, `pcy` (corresponding to parameter `obj`). The parameters for plotting are specified either as variable indexes (*integer*) or as variable names (*character*). 


<br>

#### Specification of the input parameter for plotting

Let us look on two different ways to specify parameters with the same result output (1 graph will be shown for demonstration). The first corresponds to the case, where the parameters to be plotted are specified via a *variable index* (10th variable), the second uses a *variable name* for parameter specification (non-quoted strings can be used).

```{r}
library(inLabZero)  
```

```{r echo=FALSE}
data(pcy)
```

```{r echo=FALSE}
timePlotPar(default) 
modelPlotPar(default)
```

```{r eval=FALSE}
# Plot 10th variable from the group
timePlot(pcy, 10)
```

```{r}
# Plots PARAM10 variable
timePlot(pcy, PARAM10)
```
<br> 
<br> 

It is also possible to plot a sequence of variables. In this case browsing mode is entered. For help, type `h` during browsing. 

For example (don't include the quotation marks):

<ul>
<li><strong>[enter]</strong>: next</li>
<li><strong>"5"</strong>: go next 5 steps</li>
<li><strong>"-"</strong>: go one step back</li>
<li><strong>"-3"</strong>: go 3 steps back</li>
<li><strong>"auto"</strong>: autobrowsing</li>
<li><strong>"x"</strong>: exit</li>
</ul>

```{r eval=FALSE}
# Plots 10th to 15th variable from the group
timePlot(pcy, c(10:15))
```

<br>

##### numMain

Parameter **`numMain`** enables to include/exclude index of the plotted variable in the main title. It is set to `TRUE` by default, the parameter index is excluded in this example.

```{r}
# Plots PARAM10 variable
timePlot(pcy, PARAM10, numMain=F)
```

<br> 

##### sequence

 Parameter **`sequence`** defines the `x`- axis scale. If `TRUE` points are drawn as an equidistant sequence. In the example, one point represents one lot, i.e. that the lots are sorted one by one. Otherwise the points are in natural temporal ordering and the variable specified in `timeCol` is used for the `x`- axis. By default `timeCol = pcStartDate`.


```{r}
timePlot(pcy, PARAM10, sequence=F)

```

```{r}
timePlot(pcy, PARAM10, sequence=T)
```

<br> 

The next couple of examples are divided into two sections based on whether the `z`- axis is numerical or categorical along with the usage of several other `timePlot` arguments selected based on the `z`- axis type.

<br>

#### `timePlot` argument settings for numerical `z`- axis

Parameter **`z`** specifies the variable for the `z` axis. For a numerical `z`- axis, this variable can be broken into intervals. Subsets of variable corresponding to `z` variable breaks are then plotted with different colors.

```{r}
# Select numerical z variable with input by name
timePlot(pcy, PARAM10, z="PCMYield", numMain=F)
```

<br> 

##### zBr

Parameter **`zBr`** breaks of a numeric `z` variable. Single integer value determines number of breaks. This parameter can also be specified as a vector and it can be numerical, character or POSIXct, then it represents interval breaking points. It is also possible to use special character values `min`, `mean`, `median` and `max`. 

```{r}
# Break numerical variables into intervals: keywords seq, min, max together with values
timePlotPar(legend.pos="topleft")
timePlot(pcy, PARAM10, z="PCMYield", zBr=c("min",seq(10,90,15),"max"), numMain=F)
```

<br> 

##### zBrMethod

The **`zBrMethod`** parameter defines the split methods and it works in conjunction with the `zBr` setting. The following split methods can be used:
**`median`** (breaks `z` variable with `median`), **`mean`** (breaks `z` variables with `mean`), **`thr`** (breaks `z` variable with threshold value specified by `zBr`), **`yieldHeat`** (corresponds to `zBr=c("seq(min,median,length.out=brValue)","max")`, where `brValue` is the value specified by `zBr`), **`linear`** (breaks `z` variable into equidistant intervals; number of splits is given by `zBr`), **`hist`** (breaks `z` variable using graphic’s hist function; values are passed to hist’s breaks parameter using `zBr`). The following call uses the `linear` option as the `zBrMethod`: 

```{r}
# Plot PARAM10 time series with 'linear' option as the zBrMethod
timePlot(pcy, PARAM10, z="PARAM14", zBr=10, zBrMethod="linear", numMain=F)
```

<br> 

##### zOutlier

In order to filter out outliers corresponding to the `z` variable, that would distort the real range of the variable, the `zOutlier` parameter can be applied. Parameter **`zOutlier`** is an adaptation of the `outlier` parameter for outliers filtering of the `z` variable before plotting. For more details on this parameter, refer to the `outlier` description below. 

To *combine multiple plots into one overall graph* we can use function **`par()`** and include the option **`mfrow=c(nrows, ncols)`**.

```{r fig.height=10.5}
par(mfrow=c(2,1)) 
# Plot PARAM10 time series with 'linear' option as the zBrMethod
timePlot(pcy, PARAM15, z="PARAM16", zBr=10, zBrMethod="linear", numMain=F)
# Plot PARAM10 time series with outlier filtering option
timePlot(pcy, PARAM15, z="PARAM16", zBr=10, zBrMethod="linear", zOutlier=3, numMain=F)
```

```{r echo=FALSE}
timePlotPar(default)
```

<br>

#### `timePlot` argument settings for categorical `z`- axis


Categorical `z`- axis can be broken into intervals using the **`zLevel`** parameter (*note*: In order to *include all levels*, use special value `zLevel="all"`, as depicted in the second example below). Subsets of variable corresponding to `z` variable breaks are then plotted with different colors. For a categorical `z`- axis, the `zLevel` parameter is non-null. It specifies which categories to use for split. 

```{r}
# Select categorical z variable with input by name
timePlot(pcy, PARAM10, z="center", zLevel=c(0,1), sequence=T)
```

<br> 

##### smooth

Parameter **`smooth`** adds smoothing to the chart. There are two possible options: 

* `r` - rolling function (mean by default)
* `pol`- polynomial fit


In the next example below, we will be adding rolling function (mean):


```{r}
# Set color of smoothed line (see timePlotPar section for more details)
timePlotPar(colors.line=c("#969696"), lwd=3)
# Plot rolling average
timePlot(pcy, PARAM10, z="center", zLevel="all", smooth="r", sequence=T, numMain=F)
```

<br> 

##### rWin

The **`rWin`** parameter defines the length of rolling window for rolling average charting. By default, this is set to 10, in this example, we are setting this parameter to 30.

```{r}
# Set color of smoothed line (see timePlotPar section for more details)
timePlotPar(colors.line=c("#969696"), lwd=3)
# Plot rolling average with rolling window length of 30
timePlot(pcy, PARAM10, z="center", zLevel="all", smooth="r", rWin=30, sequence=T, 
                                                                          numMain=F)
```

<br> 

##### outlier

An important part of data analysis consists in outlier filtering. Parameter **`outlier`** is applied for filtering of variable before plotting. Outliers are data points outside the interval: `<(Q1 - kl * IQR), (Q3 + kr * IQR)>`, where `Q1` and `Q3` are quartiles, `IQR` is interquartile range and `k1` and `k2` are values specified by outlier parameter as its first and second element. If single value is provided, `k1` = `k2` setting is used. No outlier filtering is applied by default. 

```{r echo=FALSE}
timePlotPar(default)
``` 

```{r}
# Select categorical z variable with input by name for all levels and outlier filtering 
timePlot(pcy, PARAM10, z="center", zLevel="all", outlier=3, sequence=T)
```

<br>  

##### yLim

There is also an option to set `y` axis limits via the **`yLim`** parameter. They represent a vector with two elements corresponding to the minimum and the maximum value of the axis. The second call in this example shows how to set the user defined `y` axis limits:

```{r fig.height=10.5}
par(mfrow=c(2,1)) 
timePlot(pcy, PARAM16, z="center", zLevel="all", numMain=F)
# User defined y axis limits
timePlot(pcy, PARAM16, z="center", zLevel="all", yLim=c(3.5,7), numMain=F)
```

<br> 

##### yLog

Setting **`yLog`** to `TRUE` will change the `y` axis to *logarithmic scale*.

```{r}
# Set y axis to logarithmic scale
timePlot(pcy, PARAM40, z="center", zLevel="all", yLog=T, sequence=T, numMain=F)
```

<br> 

##### drawOrder

The drawing order of points can be specified via the **`drawOrder`** parameter.

```{r fig.height=10.5}
par(mfrow=c(2,1)) 
# Change draw order of points
timePlot(pcy, 10, drawOrder=c(2,1))
timePlot(pcy, 10, drawOrder=c(1,2)) 
```

<br> 

Next, we will show two options for layout tuning.

<br>

##### main

The main title can be added via the **`main`** parameter.

```{r eval=FALSE}
# Add graph title
timePlot(pcy, PARAM10, z="center", zLevel=c(0,1), sequence=T, 
                                main="Center pattern generated at low PARAM10")
```
<br>

##### showLegend

To disable the displaying of the legend, set the **`showLegend`** parameter to `FALSE`.

```{r eval=FALSE}
# Hide legend
timePlot(pcy, PARAM10, z="center", zLevel="all", sequence=T, showLegend=F)
```

Let us now demonstrate the usage of the above mentioned parameters in 1 chart:

```{r}
timePlot(pcy, PARAM10, z="center", zLevel=c(0,1), sequence=T, numMain=F, showLegend=F, 
                                        main="Center pattern generated at low PARAM10")
```

<br>

#### Chart types


Parameter **`chart`** specifies chart type. We can choose from the following values:

* `p` - points (set as default)
* `r` - rolling function line (average by default)
* `pol` - polynomial fit line
* `box` - box plot

<br>

##### rFun

The **rolling function line** is shown below. The **`rFun`** is a function applied to points in the rolling window. This can take the following values: `mean` (default), `median` or `sd`. 


```{r}
# Various rolling function lines 
timePlot(pcy, PARAM10, chart="r", sequence=T, main="Rolling mean line")
```

<br> 

The next chart depicts the rolling mean line for three side IDs.  The rolling window for rolling average charting (`rWin`) is set to 20.

```{r}
timePlot(pcy, PARAM10, chart="r", rWin=20, z="siteid", zLevel=c(1,3,5), sequence=T, 
                                                            main="Rolling mean line")
```

<br> 

##### yAxisEach

In case we want to show two `y` - axis, parameter **`yAxisEach`** set to `TRUE`, will provide new `y` - axis for each variable. 



```{r}
# Use different y axes for different variables and with custom y limits
timePlot(pcy, 2, 10, chart="r", rWin=20, yAxisEach=TRUE,  yLim=c(4.6,5.2), 
                                          yLimRight=c(1700,1900), sequence=T) 
```

<br> 

##### polDegree

The **polynomial fit line** is defined via `pol` assigned to the **`chart`** parameter. The degree of the polynomial for polynomial fit charting is defined via the `polDegree` parameter (by default set to 20).

```{r}
# Plot polynomial fit line 
timePlot(pcy, PARAM10, chart="pol", sequence=T, polDegree=3, main="Polynomial fit line")
```

<br> 

Finally, the box plot chart type uses the `box` value for the **`chart`** parameter. This type of graph is suitable for smaller number of points, otherwise it could become cluttered. As a result, we will plot a subset of the dataset with `pcStartDate` after 20th of August 2016. 

```{r}
# Add box plot
timePlot(pcy[pcStartDate>"2016/08/20"], PARAM10, chart="box", sequence=T, main="Box plot")

```

<br>

#### Visualization of the model output

The limits identified by the CART model could be verified in the time plot. For more details regarding the CART model see the [CART vignette](CART.html).  

In this example, we use the value of 2 to balance the train set and select cluster `clear` for negative case sampling.  

```{r results="hide"}
 # Train CART classification on balanced data
pcyb <- model(pcy, type="class", model="cart", formula="center ~ .", cp=0.07, balance=2, 
                                                              balanceCol="clear", seed=42) 
```

Model visualizations with the use of the `modelPlot` function:

```{r}
modelPlot(pcyb, varlen=0, main="CART, cluster, classification, prune, balanced", extra=108) 
``` 

<br> 

In our analysis, we are working with wafers in the `center` cluster and we are randomly sampling twice the amount of the wafers from the cluster `clear` for balancing. In case of repeating the balancing process, random sampling will vary. 

To ensure that we use for the time plot same balanced data as for the model, we will store the selected wafers in the `usedRow` variable using the `modelTrainRow` function. The `usedRow` variable will now contain all the wafers with the `center` cluster and samples from the cluster `clear`.  

```{r}
usedRow <- modelTrainRow(pcyb)
```

Since we have stored all the wafers in the `center` cluster and samples from cluster `clear` in the `usedRow` variable, the `timePlot` can be called on `pcyb[usedRow]` subset so that only the data used in the model is displayed. The CART limits are added using `abline` function, a red horizontal line at the value of 1713.  


```{r}
timePlotPar(legend.pos="bottomleft")
 # Select categorical z variable with input by name for all levels
timePlot(pcyb[usedRow], PARAM10, z="center", zLevel="all", numMain=F)
# add CART limit - red horizontal line
abline(h=1713, col="red", lty=3)  
```

<br>

### 2. Time plot visual parameters - `timePlotPar` {#section2}  

The **`timePlotPar`** function enables to get, set and reload default visual parameters of the `timePlot`. For a detailed list of all parameters that are available, see `?timePlotPar`. Use parameter names to get values, use assignment to set parameter values and use keyword `default` to load default values. Let us look at some of the most used parameters and their usage. In order to see the effect on the `timePlot`, the `timePlotPar` needs to be run before the `timePlot` is plotted. 


<br>

##### colors.alpha

To get the value of a visual parameter, in our case value of color transparency setting per chart type, use: 

```{r}
# Get value of color transparency setting per chart type
timePlotPar(colors.alpha)
```

<br>

##### legend.pos

The *legend position* is accessible via the **`legend.pos`** parameter. In the call below, the legend position is set to top left: 

```{r}
# Set legend position to top left     
timePlotPar(legend.pos="topleft")
timePlot(pcy, PARAM10, z="center", zLevel=c(0,1), numMain=F) 
```

<br> 

##### colors.zLevel

*Color scheme* for categorical variables can be tuned via the **`colors.zLevel`** parameter. Please refer to [color palette](https://colorbrewer2.org/#type=qualitative&scheme=Set3&n=12) for more details on color specification.  

```{r}
# Set color scheme to user defined values
timePlotPar(colors.zLevel=c("#8dd3c7","#fdb462"), legend.pos="topright")
timePlot(pcy, PARAM10, z="center", zLevel=c(0,1), numMain=F) 
```

```{r echo=FALSE}
timePlotPar(default)
```

<br> 

##### colors.line

Parameter **colors.line** represents the *colors of smoothed line*, while the **`lwd`** defines its width.  

```{r}
# Set the color and width of the smoothed line
timePlotPar(colors.line=c("#969696"), lwd=3)
timePlot(pcy, PARAM10, z="center", zLevel="all", smooth="r", rWin=30, sequence=T, 
                                                                        numMain=F)
```


```{r echo=FALSE}
timePlotPar(default)
```

<br> 

##### cexAxis

Parameter **`cexAxis`** is used to set the size of the axis annotation. 

```{r}
timePlotPar(cexAxis=1.2)
timePlot(pcy, PARAM10, z="center", zLevel=c(0,1), numMain=F)
```

<br> 

In case that we have changed the visual parameters and we want to reload the default settings, the following call can be applied: 

```{r}
# Reload default visual parameters
timePlotPar(default) 
```




